---
title: "Simulation of CER Method for mixed binary/continuous endpoints"
output-dir: output
author: B. Fallmann
date: today
email: "benjamin.fallmann\\@meduniwien.ac.at"
format:
  html:
    embed-resources: true
    fig-format: svg
    title-block-banner: true
    toc: true
    toc-location: left
    papersize: a4
  pdf: 
    fig-pos: 'H'
    documentclass: article
execute:
  freeze: true
  echo: false 
  message: false 
  warning: false
number-sections: true
fig-width: 6
fig-asp: 0.618
out-width: "70%"
---

```{r}
library(targets)
library(gt)
```

A simulation of the CER method for mixed continuous and binary endpoints was
conducted.
We assumed a trial with 4 different treatment arms, each compared pair-wise to
a common control.
For each arm, we tested two hypotheses, one primary and one secondary.
The graph used for describing the testing strategy can be found in @fig-graph.

The testing strategy for the intersection hypotheses used parametric tests for adjusting among p-values with
known correlation (that is hypothesis sharing the same control) and Bonferroni
otherwise.

```{r}
#| label: fig-graph
#| fig-cap: Graph of the multiple comparison strategy used for the simulation.
#|   Here hypotheses H1 - H4 are the primary endpoints for arms 1 to 4, and H5
#|  - H8 are the corresponding secondary arms.
library(graphicalMCP)
tar_load(design)
graph <- graph_create(design$weights, design$test_m)

plot_layout <- rbind(
  c(0, 0.5),
  c(0.3, 0.5),
  c(0.5, 0.5),
  c(0.8, 0.5),
  c(0, 0),
  c(0.3, 0),
  c(0.5, 0),
  c(0.8, 0)
)
plot(graph, layout = plot_layout)
```

The primary endpoints were continuous outcomes, always simulated as normal
variables with a variance of 1.
Whenever a endpoint was assumed to be effective, we assumed a effect size of
0.4, that is the expected difference between control and treatment.
The secondary endpoints were binary outcomes. The control always had a response
rate of 0.1, and the treatment arm had the same response rate of 0.1 for
ineffective arms and a response rate of 0.25 for effective arms.

To simulate a correlation between related primary and secondary arm, data was
generated in the following way:
Assume a fixed correlation $c$. We draw data from a bivariate normal
distribution with the given correlation and covariance 1.
The first part is used as the outcome for the primary endpoint. The second
endpoint is set to 1 whenever the second part of the draw is below the
corresponding quantile of the desired response rate, and to 0 otherwise.

We used a sample size of 100 subjects for each arm, and did an interim analysis
with possibility of adjustment and early rejection with the data of the first 50
subjects.

For the continuous endpoints, p-values were generated using one-sided t-tests
with degrees of freedom under assuption of equal variance. For the binary
endpoints, a one-sided two-proportion z-test with pooled variance was used.

For the interim adaptions, we considered the following adaption rulesets:

* Conservative: Drop arm if p-value of primary hypotheses greater or equal 0.75
* Normal: Drop arm if p-value of primary hypotheses greater or equal 0.5
* Agressive: Drop arm if p-value of primary hypotheses greater or equal 0.25
* Ultra Agressive: Keep only arm with lowest primary hypotheses p-value, and stop
  if the corresponding arm (i.e. primary and secondary hypothesis) get rejected
  at the interim

Whenever an arm was dropped, all future patients that would have been assigned
to that arm were reassigned to the remaining arms and control group equally.

For the power analysis, each of those strategies was simulated for the first 1,
2, 3 or all hypotheses being effective, while assuming a correlation $c$ of 0.5.
For the analysis of family-wise error control (FWER), each of the strategies was
tested under correlation of 0.3, 0.5, and 0.8 respectively.

For each scenario, we conducted 100,000 first stage simulations. For each
simulated first stage, 100 stage two simulations were conducted.
This has the benefit of being more computationally efficient compared to
simulating each trial once, since the main computational effort is on adjusting the
bounds for rejection after the adaptions from the interim analysis.

```{r}
#| label: tbl-power-res
#| tbl-cap: Table describing the results for the different scenarios for the
#|   power analysis. Disjunctive refers to the disjunctive power, so the chance
#|   that any effective hypotheses is rejected. Conjunctive power is the chance
#|   that all effective hypotheses are rejected. This is therefore not expected
#|   to happen in the ultra agressive scenario if more than one arm is
#|   effective. CI is the 95% confidence interval
tar_load(power_tbl)
power_tbl$tbl_bin |>
  tab_options(table.font.size = 11)
```

```{r}
#| label: tbl-fwer_res
#| tbl-cap: Table describing the results for the different scenarios for the
#|   analysis of the FWER.
#|   CI is the 95% confidence interval
tar_load(fwer_tbl)
fwer_tbl$tbl_bin |>
  tab_options(table.font.size = 11)
```

```{r}
#The main difference to the scenarios with all continuous outcomes is that the
#conjunctive power is much lower, especially in scenarios with many effective
#arms. This is to be expected, since the non-centrality parameter for a single
#hypothesis of continuous outcomes is
#$$\frac{0.4}{2 / \sqrt{100}} = 2,$$
#versus the parameter for the binary outcomes is
#$$\frac{0.15}{\sqrt{0.15 \cdot 0.85 \cdot \frac{2}{100}}} \approx 2.73.$$
```
