---
title: "Simulation of CER Method for mixed binary/continuous endpoints"
output-dir: output
author: B. Fallmann
date: today
email: "benjamin.fallmann\\@meduniwien.ac.at"
format:
  html:
    embed-resources: true
    fig-format: svg
    title-block-banner: true
    toc: true
    toc-location: left
    papersize: a4
  pdf: 
    fig-pos: 'H'
    documentclass: article
execute:
  freeze: true
  echo: false 
  message: false 
  warning: false
number-sections: true
fig-width: 6
fig-asp: 0.618
out-width: "70%"
---

```{r}
library(targets)
library(gt)
```

A simulation of the CER method for mixed continuous and binary endpoints was
conducted.
We assumed a trial with 4 different treatment arms, each compared pair-wise to
a common control.
For each arm, we tested two hypotheses, one primary and one secondary.
The graph used for describing the testing strategy can be found in @fig-graph.

The testing strategy for the intersection hypotheses used parametric tests for adjusting among p-values with
known correlation (that is hypothesis sharing the same control) and Bonferroni
otherwise.

```{r}
#| label: fig-graph
#| fig-cap: Graph of the multiple comparison strategy used for the simulation.
#|   Here hypotheses H1 - H4 are the primary endpoints for arms 1 to 4, and H5
#|  - H8 are the corresponding secondary arms.
library(graphicalMCP)
tar_load(design)
graph <- graph_create(design$weights, design$test_m)

plot_layout <- rbind(
  c(0, 0.5),
  c(0.3, 0.5),
  c(0.5, 0.5),
  c(0.8, 0.5),
  c(0, 0),
  c(0.3, 0),
  c(0.5, 0),
  c(0.8, 0)
)
plot(graph, layout = plot_layout)
```


We used a sample size of 100 subjects for each arm, and did an interim analysis
with possibility of adjustment and early rejection with the data of the first 50
subjects.


For the interim adaptions, we considered the following adaption rulesets:

* Conservative: Drop arm if p-value of primary hypotheses greater or equal 0.75
* Normal: Drop arm if p-value of primary hypotheses greater or equal 0.5
* Agressive: Drop arm if p-value of primary hypotheses greater or equal 0.25
* Ultra Agressive: Keep only arm with lowest primary hypotheses p-value, and stop
  if the corresponding arm (i.e. primary and secondary hypothesis) get rejected
  at the interim

Whenever an arm was dropped, all future patients that would have been assigned
to that arm were reassigned to the remaining arms and control group equally.
For the power analysis, each of the adaption strategies was simulated for the last 1,
2, 3 or all hypotheses being effective.

For each scenario, we conducted 100,000 first stage simulations. For each
simulated first stage, 100 stage two simulations were conducted.
This has the benefit of being more computationally efficient compared to
simulating each trial once, since the main computational effort is on adjusting the
bounds for rejection after the adaptions from the interim analysis.

# Continuous Outcomes

In these scenarios, all outcomes are generated as continous normal variables.
The different arms and the control group were identically and independently distributed.
In each arm, a correlation between the primary and secondary endpoint was
assumed.
For the power analysis, all scenarios assumed a correlation $c$ of 0.5.
For the analysis of family-wise error control (FWER), each of the adaption strategies was
tested under correlation of 0.3, 0.5, and 0.8 respectively.

All arms had a variance of 1.
The mean was taken as 0 for the control and all ineffective arms, while
any effective arm had an mean of 0.4.

P-values were generated using one-sided t-tests
with degrees of freedom under assuption of equal variance. 

```{r}
#| label: tbl-power-res
#| tbl-cap: Table describing the results for the different scenarios for the
#|   power analysis und continous endpoints. Disjunctive refers to the disjunctive power, so the chance
#|   that any effective hypotheses is rejected. Conjunctive power is the chance
#|   that all effective hypotheses are rejected. This is therefore not expected
#|   to happen in the ultra agressive scenario if more than one arm is
#|   effective. CI is the 95% confidence interval.
tar_load(power_tbl)
power_tbl$tbl_cont |>
  tab_options(table.font.size = 11)
```

```{r}
#| label: tbl-fwer_res
#| tbl-cap: Table describing the results for the different scenarios for the
#|   analysis of the FWER under continous endpoints.
#|   CI is the 95% confidence interval.
tar_load(fwer_tbl)
fwer_tbl$tbl_cont |>
  tab_options(table.font.size = 11)
```

# Mixed Continuous and Binary Outcomes

The primary endpoints were continuous outcomes were as in the continuous
scenarios, that is always simulated as normal
variables with a variance of 1.
Whenever a endpoint was assumed to be effective, we assumed a effect size of
0.4, that is the expected difference between control and treatment.
The secondary endpoints were binary outcomes. The control always had a response
rate of 0.1, and the treatment arm had the same response rate of 0.1 for
ineffective arms and a response rate of 0.25 for effective arms.

To simulate a correlation between related primary and secondary arm, data was
generated in the following way:
Assume a fixed correlation $c$. We draw data from a bivariate normal
distribution with the given correlation and covariance 1.
The first part is used as the outcome for the primary endpoint. The second
endpoint is set to 1 whenever the second part of the draw is below the
corresponding quantile of the desired response rate, and to 0 otherwise.

For the binary
endpoints, a one-sided .two-proportion z-test with pooled variance was used.
For the continous arms we used one-sided t-test with equal variance as before.

```{r}
#| label: tbl-power-res-bin
#| tbl-cap: Table describing the results for the different scenarios for the
#|   power analysis under mixed endpoints. Disjunctive refers to the disjunctive power, so the chance
#|   that any effective hypotheses is rejected. Conjunctive power is the chance
#|   that all effective hypotheses are rejected. This is therefore not expected
#|   to happen in the ultra agressive scenario if more than one arm is
#|   effective. CI is the 95% confidence interval.
power_tbl$tbl_bin |>
  tab_options(table.font.size = 11)
```

```{r}
#| label: tbl-fwer-res-bin
#| tbl-cap: Table describing the results for the different scenarios for the
#|   analysis of the FWER under mixed enpoints.
#|   CI is the 95% confidence interval.
fwer_tbl$tbl_bin |>
  tab_options(table.font.size = 11)
```
